### 2016-07-05 introduction.tex
typo fixes

Recent results on reasoning about application invariants in the
presence of weak \emph{consistency}~\cite{burckhardt14, redblueosdi,
  redblueatc, ecinec, gotsmanpopl16} address broadly related concerns.
At a high-level, reasoning about weak consistency involves reasoning
about the concurrent execution of strongly atomic and fully isolated
operations against different (potentially replicated), but consistent
states of the database or data store, where each operation locally
preserves application invariants. In contrast, reasoning about weak
isolation involves reasoning about the concurrent execution of
transactions, whose operations witness different states which,
although consistent individually, may not be obviously reconciled into
a consistent global view.  Intuitively, the latter problem reduces to
the former if all transactions contain a single
operation. Furthermore, weak isolation can exist independent of weak
consistency, as evident from the presence of weakly isolated
transactions on sequentially consistent relational databases; such
systems existed long before the advent of weakly consistent NoSQL
stores.  Thus, frameworks for reasoning about weak isolation will
necessarily have to generalize the reasoning frameworks developed for
weak consistency in new and important ways.

@@

In this paper, we propose a program logic for weakly-isolated
transactions that realizes this goal.  In particular, we develop a set
of syntax-directed compositional proof rules that allow developers to
build correctness proofs for transactional programs in the presence of
a weakly isolated concurrency control mechanism.  A key novelty of our
approach is that it is parametric over the \emph{isolation semantics}
of transactions in the program, as well as the \emph{consistency
  semantics} of the underlying store. In concrete terms, this means
that, unlike recent work focused on formalizing the semantics of weak
consistency~\cite{gotsmanpopl16, redblueatc, ecinec}, our system model
does not assume a minimum or predefined set of consistency or
isolation levels. Instead, our operational semantics admits
declarative specifications of transaction isolation and store
consistency levels (\eg~\cite{pldi15,gotsmanconcur15}), and generates
executions that are guaranteed to satisfy these specifications. The
result is a flexible system model that is general enough to
incorporate the semantics of a range of isolation levels on a variety
of stores (\eg the sequentially consistent store of~\cite{adyaphd}, or
the causally consistent store of~\cite{gotsmanpopl16}, etc).  Our key
technical contribution is thus a set of proof rules that demonstrate
that a program with a given selection of isolation levels for its
transactions preserves its invariants when executed on a store
equipped with certain consistency guarantees (e.g., causal
visibility).

The paper makes the following contributions:
\begin{itemize}
  \item We develop a semantics for a core language equipped with
    weakly-isolated transactions, demonstrating that a general
    parametric isolation semantics can be expressed as
    well-formedness constraints on a program's execution.
  \item We propose a compositional proof system for this language
    capable of relating high-level application invariants to the structure
    of traces induced by the operational semantics, in which transactions
    are associated with specific weak isolation levels.
  \item We define a \emph{maximum visibility principle} to reconcile
    store consistency guarantees with transaction isolation
    requirements, and use it to instantiate our operational model and
    the proof system for various kinds of (weakly-consistent) data
    stores.
  \item Case studies modeled after real-world scenarios demonstrate
    the applicability and utility of our proof methodology.
\end{itemize}

\noindent Our results provide the first (to the best of our
knowledge) mechanism that precisely relates high-level program
invariants to low-level weak isolation and store consistency
guarantees, thereby allowing weakly-isolated transactions to enjoy the
same rigorous reasoning capabilities as their strongly-isolated
(serializable) counterparts.

>>>>

Recent results on reasoning about application invariants in the
presence of weak \emph{consistency}~\cite{burckhardt14, redblueosdi,
  redblueatc, ecinec, gotsmanpopl16} address broadly related concerns.
At a high-level, reasoning about weak consistency involves reasoning
about the concurrent execution of strongly atomic and fully isolated
operations against different (potentially replicated), but consistent
states of the database or data store, where each operation locally
preserves application invariants. In contrast, reasoning about weak
isolation involves reasoning about the concurrent execution of
transactions, whose operations witness different states which,
although consistent individually, may not be obviously reconciled into
a consistent global view.  Intuitively, the latter problem reduces to
the former if all transactions contain a single
operation. Furthermore, weak isolation can exist independent of weak
consistency, as evident from the presence of weakly-isolated
transactions on sequentially consistent relational databases; such
systems existed long before the advent of weakly consistent NoSQL
stores.  Thus, frameworks for reasoning about weak isolation will
necessarily have to generalize the reasoning frameworks developed for
weak consistency in new and important ways.

@@

In this paper, we propose a program logic for weakly-isolated
transactions that realizes this goal.  In particular, we develop a set
of syntax-directed compositional proof rules that allow developers to
build correctness proofs for transactional programs in the presence of
a weakly-isolated concurrency control mechanism.  A key novelty of our
approach is that it is parametric over the \emph{isolation semantics}
of transactions in the program, as well as the \emph{consistency
  semantics} of the underlying store. In concrete terms, this means
that, unlike recent work focused on formalizing the semantics of weak
consistency~\cite{gotsmanpopl16, redblueatc, ecinec}, our system model
does not assume a minimum or predefined set of consistency or
isolation levels. Instead, our operational semantics admits
declarative specifications of transaction isolation and store
consistency levels (\eg~\cite{pldi15,gotsmanconcur15}), and generates
executions that are guaranteed to satisfy these specifications. The
result is a flexible system model that is general enough to
incorporate the semantics of a range of isolation levels on a variety
of stores (\eg the sequentially consistent store of~\cite{adyaphd}, or
the causally consistent store of~\cite{gotsmanpopl16}, etc).  Our key
technical contribution is thus a set of proof rules that demonstrate
that a program with a given selection of isolation levels for its
transactions preserves its invariants when executed on a store
equipped with certain consistency guarantees (e.g., causal
visibility).

The paper makes the following contributions:
\begin{itemize}
  \item We develop a semantics for a core language equipped with
    weakly-isolated transactions, demonstrating that a general
    parametric isolation semantics can be expressed as
    well-formedness constraints on a program's execution.
  \item We present a compositional proof system for this language
    capable of relating high-level application invariants to the structure
    of traces induced by the operational semantics, in which transactions
    are associated with specific weak isolation levels.
  \item We define a \emph{maximum visibility principle} to reconcile
    store consistency guarantees with transaction isolation
    requirements, and use it to instantiate our operational model and
    the proof system for various kinds of (weakly-consistent) data
    stores.
  \item Case studies modeled after real-world scenarios demonstrate
    the applicability and utility of our proof methodology.
\end{itemize}

\noindent Our results provide the first (to the best of our knowledge)
mechanism that precisely and uniformly relates high-level program
invariants to low-level weak isolation and store consistency
guarantees, thereby allowing weakly-isolated transactions to enjoy the
same rigorous reasoning capabilities as their strongly-isolated
(serializable) counterparts.
